---
title: "P8105 Homework 2 [td2704]"
output: github_document
---


```{r, echo=FALSE, message=FALSE}

library(tidyverse)
library(readxl)

```

## Problem 1

Mr. Trash Wheel sheet.

```{r}

# order below
## specify sheet, omit non-data entries
## clean variable names to standardize
## omit rows without dumpster-specific data (remove na's in dumpster column)
## round number of sports balls to nearest integer

trashwheel_df = 
  read_excel("data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
                           sheet = "Mr. Trash Wheel",
                           range = "A2:N408") %>%
  janitor::clean_names() %>%
  drop_na(dumpster) %>%
  mutate(
    sports_balls = round(sports_balls, digits = 0)
  )

trashwheel_df

```

Precipitation data for 2018 annd 2019.

```{r}

precip18_df = 
  read_excel("data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
             sheet = "2018 Precipitation",
             range = "A2:B14") %>%
  janitor::clean_names() %>%
  mutate(
    year = rep(2018, times = 12)
  ) %>%
  relocate(year)

precip18_df

```

```{r}

precip19_df = 
  read_excel("data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
             sheet = "2019 Precipitation",
             range = "A2:B14") %>%
  janitor::clean_names() %>%
  mutate(
    year = rep(2019, times = 12)
  ) %>%
  relocate(year)

precip19_df

```

```{r}

precip_combined_df = 
  bind_rows(precip18_df,precip19_df) %>%
  mutate(
    month = month.name[month]
  )

precip_combined_df

```

The Mr. Trash Wheel dataset has `r nrow(trashwheel_df)` rows of 
`r length(colnames(trashwheel_df))` variables: `r colnames(trashwheel_df)`.
It tracks homes powered by various recycled/reused items over the years
`r unique(pull(trashwheel_df,year))`. The median number of sports balls in a 
dumpster in 2019 was `r median(pull(filter(trashwheel_df,year==2019),sports_balls),na.rm = TRUE)`.
The resulting precipitation dataset has `r nrow(precip_combined_df)` rows of 
`r length(colnames(precip_combined_df))` variables: `r colnames(precip_combined_df)`.
Years included in this dataset are `r unique(pull(precip_combined_df,year))`,
and the amount of precipitation is listed for each month, year pair.
From this data, the total precipitation in 2018 was 
`r sum(pull(filter(precip_combined_df,year==2018),total), na.rm = TRUE)`.







