---
title: "P8105 Homework 2 [td2704]"
output: github_document
---


```{r, echo=FALSE, message=FALSE}

library(tidyverse)
library(readxl)
library(lubridate)
library(dplyr)

```

## Problem 1

Mr. Trash Wheel sheet.

```{r}

# order below
## specify sheet, omit non-data entries
## clean variable names to standardize
## omit rows without dumpster-specific data (remove na's in dumpster column)
## round number of sports balls to nearest integer

trashwheel_df = 
  read_excel("data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
                           sheet = "Mr. Trash Wheel",
                           range = "A2:N408") %>%
  janitor::clean_names() %>%
  drop_na(dumpster) %>%
  mutate(
    sports_balls = round(sports_balls, digits = 0)
  )

trashwheel_df

```

Precipitation data for 2018 annd 2019.

```{r}

precip18_df = 
  read_excel("data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
             sheet = "2018 Precipitation",
             range = "A2:B14") %>%
  janitor::clean_names() %>%
  mutate(
    year = rep(2018, times = 12)
  ) %>%
  relocate(year)

precip18_df

```

```{r}

precip19_df = 
  read_excel("data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
             sheet = "2019 Precipitation",
             range = "A2:B14") %>%
  janitor::clean_names() %>%
  mutate(
    year = rep(2019, times = 12)
  ) %>%
  relocate(year)

precip19_df

```

```{r}

precip_combined_df = 
  bind_rows(precip18_df,precip19_df) %>%
  mutate(
    month = month.name[month]
  )

precip_combined_df

```

The Mr. Trash Wheel dataset has `r nrow(trashwheel_df)` rows of 
`r length(colnames(trashwheel_df))` variables: `r colnames(trashwheel_df)`.
It tracks homes powered by various recycled/reused items over the years
`r unique(pull(trashwheel_df,year))`. The median number of sports balls in a 
dumpster in 2019 was `r median(pull(filter(trashwheel_df,year==2019),sports_balls),na.rm = TRUE)`.
The resulting precipitation dataset has `r nrow(precip_combined_df)` rows of 
`r length(colnames(precip_combined_df))` variables: `r colnames(precip_combined_df)`.
Years included in this dataset are `r unique(pull(precip_combined_df,year))`,
and the amount of precipitation is listed for each month, year pair.
From this data, the total precipitation in 2018 was 
`r sum(pull(filter(precip_combined_df,year==2018),total), na.rm = TRUE)`.


## Problem 2

Assume any nonzero values count as TRUE for president column

```{r}

pols_month_df2 = 
  read_csv("data/fivethirtyeight_datasets/pols-month.csv") %>%
  janitor::clean_names() %>%
  filter(
    prez_gop == 2
  )
pols_month_df2

```


```{r}

pols_month_df = 
  read_csv("data/fivethirtyeight_datasets/pols-month.csv") %>%
  janitor::clean_names() %>%
  separate(mon, into = c("year", "month", "day"), sep = "-", convert = TRUE) %>%
  mutate(
    month = month.name[month]
  ) %>%
  pivot_longer(
    cols = starts_with("prez_"),
    names_to="president",
    names_prefix = "prez_",
    values_to="president_boolean"
  ) %>%
  filter(president_boolean!=0) %>%
  select(-president_boolean, -day)

pols_month_df
```



```{r}

snp_df = 
  read_csv("data/fivethirtyeight_datasets/snp.csv") %>%
  janitor::clean_names() %>%
  separate(date, into = c("month", "day", "year"), sep = "/", convert = TRUE) %>%
  relocate(year,month) %>%
  mutate(
    year = ifelse(year >= 1950 %% 100, 1900+year, 2000+year), # to get proper 4 digit year
  ) %>%
  select(-day) %>%
  arrange(year,month) %>%
  mutate(
    month = month.name[month]
  )


snp_df

```


```{r}

unemployment_df = 
  read_csv("data/fivethirtyeight_datasets/unemployment.csv") %>%
  janitor::clean_names() %>%
  pivot_longer(
    jan:dec,
    names_to="month",
    values_to="perc_unemployment"
  ) %>%
  mutate(
    year = as.integer(year),
    month = month.name[(match(month, tolower(month.abb)))]
  )

unemployment_df

```

Merge snp into pols first. Then merge unemployment into result.

Inner join:

```{r}

joined_df = 
  inner_join(pols_month_df, snp_df, by = c("year","month")) %>%
  inner_join(unemployment_df, by = c("year","month"))

joined_df

```


The first dataset pols_month_df contains the number of national politicians who
are Democratic or Republican at any given time, including president, governors,
senators, etc - it has `r nrow(pols_month_df)` rows with variables `r colnames(pols_month_df)`.
This dataset comes from the years `r min(pull(pols_month_df,year))` to `r max(pull(pols_month_df,year))`.
The second dataset snp_df contains the S&P stock market index over year, month pairs
- it has `r nrow(snp_df)` rows with variables `r colnames(snp_df)`.
This dataset comes from the years `r min(pull(snp_df,year))` to `r max(pull(snp_df,year))`.
The final dataset unemployment_df contains the unemployment percentage in the US
over year, month pairs - it has `r nrow(unemployment_df)` rows with variables 
`r colnames(unemployment_df)`. This dataset comes from the years 
`r min(pull(unemployment_df,year))` to `r max(pull(unemployment_df,year))`. 
The resulting dataset joined_df combines all this data into one dataframe of size 
`r nrow(joined_df)` rows and `r ncol(joined_df)` columns with variables `r colnames(joined_df)`.
It includes, by month,year pair, the number of Democratic and Republican 
politicians across the US, as well as  the S&P stock index and unemployment rate
at that time as well to provide pointers for the economy at that time. This
joined dataset spans the years `r min(pull(joined_df,year))` to `r max(pull(joined_df,year))`.
This joined dataset has only year,month pairs data that exists in all 3  of the
individual datasets.

If we use left join instead:

Left join:

```{r}

joined_df_left = 
  left_join(pols_month_df, snp_df, by = c("year","month")) %>%
  left_join(unemployment_df, by = c("year","month"))

joined_df_left

```
The first dataset pols_month_df contains the number of national politicians who
are Democratic or Republican at any given time, including president, governors,
senators, etc - it has `r nrow(pols_month_df)` rows with variables `r colnames(pols_month_df)`.
This dataset comes from the years `r min(pull(pols_month_df,year))` to `r max(pull(pols_month_df,year))`.
The second dataset snp_df contains the S&P stock market index over year, month pairs
- it has `r nrow(snp_df)` rows with variables `r colnames(snp_df)`.
This dataset comes from the years `r min(pull(snp_df,year))` to `r max(pull(snp_df,year))`.
The final dataset unemployment_df contains the unemployment percentage in the US
over year, month pairs - it has `r nrow(unemployment_df)` rows with variables 
`r colnames(unemployment_df)`. This dataset comes from the years 
`r min(pull(unemployment_df,year))` to `r max(pull(unemployment_df,year))`. 
The resulting dataset left_joined_df combines all this data into one dataframe of size 
`r nrow(joined_df_left)` rows and `r ncol(joined_df_left)` columns with variables `r colnames(joined_df_left)`.
It includes, by month,year pair, the number of Democratic and Republican 
politicians across the US, as well as  the S&P stock index and unemployment rate
at that time as well to provide pointers for the economy at that time. This
joined dataset spans the years `r min(pull(joined_df_left,year))` to `r max(pull(joined_df_left,year))`.
This left joined dataset has all year,month pairs data that exist in pols_month_df,
where some values for close and unemployment may be N/A if those year,month pairs
didn't exist in the other datasets. Hence, this is a larger dataset than the
inner joined dataframe but may be more incomplete.


## Problem 3

Tidy the data.

```{r}

babynames_df = 
  read_csv("data/Popular_Baby_Names.csv") %>%
  janitor::clean_names()  %>%
  mutate(
    ethnicity = replace(ethnicity, ethnicity == "WHITE NON HISP", "WHITE NON HISPANIC"),
    ethnicity = replace(ethnicity, ethnicity == "ASIAN AND PACI", "ASIAN AND PACIFIC ISLANDER"),
    ethnicity = replace(ethnicity, ethnicity == "BLACK NON HISP", "BLACK NON HISPANIC"),
    childs_first_name = toupper(childs_first_name)
  ) %>%
  distinct(.keep_all=TRUE)

babynames_df
```


rank in popularity of the name “Olivia” as a female baby name over time:


```{r}

olivia_rank = 
  filter(babynames_df,childs_first_name=="OLIVIA") %>%
  filter(gender=="FEMALE") %>%
  arrange(year_of_birth) %>%
  select(-count) %>%
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "rank"
  ) %>%
  relocate(childs_first_name) %>%
  knitr::kable()
  

olivia_rank

```


most popular name among male children over time

```{r}

male_rank = 
  filter(babynames_df,gender=="MALE") %>%
  arrange(rank,year_of_birth) %>%
  select(-count) %>%
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "rank"
  ) %>%
  relocate(childs_first_name) %>%
  knitr::kable()
  

male_rank
```


Finally, for male, white non-hispanic children born in 2016, produce a scatter plot showing the number of children with a name (y axis) against the rank in popularity of that name (x axis).

```{r}

male_whitenonhisp_2016_df = 
  filter(babynames_df, gender=="MALE") %>%
  filter(ethnicity=="WHITE NON HISPANIC") %>%
  filter(year_of_birth=='2016')

male_whitenonhisp_2016_df

```

```{r}

male_whitenonhisp_2016_df %>%
  ggplot(aes(x=rank,y=count)) +
  geom_point() +
  labs(
    title = "number of children with a name against the rank in popularity of that name"
  )

```

